Structural replication of nanoporous media using procedural noise

### 2017-02-06 ms.tex
Fixes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate geometries at arbitrary scales. In addition, this method is extremely fast and has low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media. For a given data set and a model, we wish to estimate the parameters that best represent the data. This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system. The simulated system is confined to a cube of size $\SI{159}{\angstrom^{3}}$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.

@@

\subsection{Model verification: Parameter estimation of mock data}
\begin{figure}[htb!]
\includegraphics[width=.45\textwidth]{model_test.png}
\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
\label{fig:mockdata}
\end{figure}
In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values. In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$. We simulate a data set of $159 ^3 \AA$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$. The resulting 3D structure is depicted in figure \ref{fig:mockdata}.

@@

\subsection{Parameter estimation of simulated nanoporous silica}
%\begin{figure}
%\includegraphics[width=.49\textwidth]{sio_porous.png}
%\caption{Simulated silica created with LAMMPS. The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved. Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
%\label{fig:si2data}
%\end{figure}
\begin{figure*}
\includegraphics[width=.99\textwidth]{results_porous_full.png}
\caption{Results from the analysis of a simulated $159^3$ \AA SiO$_2$ nanoporous data set. The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors. Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
\label{fig:porous_results1}
\end{figure*}

@@

The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}
In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise. We have developed tools for creating such porous materials, and also estimating model parameters given data. We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model. In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data. However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
In this paper, we aim to develop a method for producing statistically correct porous media using noise functions. In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall. By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales. It is therefore possible to generate geometries at arbitrary scales. In addition, this method is extremely fast and has low memory footprints, requiring very little computational power. After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g. molecular dynamics) or with our method. The radial distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model. Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters. Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}

@@

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media. For a given data set and a model, we wish to estimate the parameters that best represent the data. This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system. The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.

@@

\subsection{Model verification: Parameter estimation of mock data}
\begin{figure}[htb!]
\includegraphics[width=.45\textwidth]{model_test.png}
\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
\label{fig:mockdata}
\end{figure}
In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values. In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$. We simulate a data set of $\SI{159} {\angstrom}^3$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$. The resulting 3D structure is depicted in figure \ref{fig:mockdata}.

@@

\subsection{Parameter estimation of simulated nanoporous silica}
%\begin{figure}
%\includegraphics[width=.49\textwidth]{sio_porous.png}
%\caption{Simulated silica created with LAMMPS. The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved. Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
%\label{fig:si2data}
%\end{figure}
\begin{figure*}
\includegraphics[width=.99\textwidth]{results_porous_full.png}
\caption{
Results from the analysis of a simulated $\SI{159} {\angstrom}^3$ SiO$_2$ nanoporous data set. The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors. Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.
}

\label{fig:porous_results1}
\end{figure*}

@@

The porosity of the input SiO$_2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}
In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise. We have developed tools for creating such porous materials, and also estimating model parameters given data. We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model. In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data. However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
